---
title: "Highschool Test Scores"
author: "Matthew Hondrakis"
output: 
  html_document:
    toc: true
    toc_float: true
date: "`r format(Sys.time(), '%d %B %Y')`"
editor_options: 
  chunk_output_type: console
---

# Student Test Scores


```{r, include=FALSE}
library(tidyverse)
library(GGally)
library(plotly)
theme_set(theme_bw())
stest <- read_csv("StudentsPerformance.csv", col_types = "fffffddd")
stest <- stest %>%
  rename(`parent ed` =`parental level of education`,
         `test prep` = `test preparation course`,
         `math` = `math score`,
         `reading` = `reading score`,
         `writing` = `writing score`)
stest$`parent ed` <- factor(stest$`parent ed`, 
                            levels = c("some high school",
                                       "high school",
                                       "some college",
                                       "associate's degree",
                                       "bachelor's degree",
                                       "master's degree"))
levels(stest$gender) <- list(A = "female", B = "male")
stest$`race/ethnicity` <- factor(stest$`race/ethnicity`, 
                                 levels = c("group A", "group B", 
                                            "group C","group D", "group E"))
```

## Structure of Data Set
```{r Data, message=FALSE, warning=FALSE}
str(stest)
```

## Correlation Plot

```{r, warning=FALSE, message=FALSE}
stest %>%
  select(math, writing, reading) %>%
  ggpairs()
```


There is a clear correlation between the pair of scores. With the highest **correlation** coefficient of `r cor(stest$"writing", stest$"reading")` being between *Reading* and *Writing*.


## By Gender

### Math

```{r fig.show = "hold", message=FALSE, warning=FALSE}
gplot <- function(data = stest, x, color){
  ggplot(data, aes({{x}}, color = {{color}}))
}
gplot(stest, math, gender) + geom_density()

gplot(stest, math, gender) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                         axis.ticks.y = element_blank())
```


The variance is approximately the same and they are normally distributed. Thus, We can conduct a **t-test** to evaluate whether the difference in means between the two groups is statistically significant.


```{r message=FALSE, warning=FALSE}
t.test(stest$`math`~stest$gender, conf.level = 0.99)
```

With 99% confidence, we can assume the difference in means between the two genders is between (-7.52,-2.66).

### Writing

```{r fig.show = "hold", message=FALSE, warning=FALSE}
gplot(stest,writing, gender) + geom_density()

gplot(stest, writing, gender) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                       axis.ticks.y = element_blank())
```


It appears that in the case of writing, it is *gender A* that scores better than *gender B*. 


```{r message=FALSE, warning=FALSE}
t.test(stest$`writing`~stest$gender, conf.level = 0.99)
```

As we expected, with 99% confidence, we can assume the difference in means between the two genders is between (6.79,11.52). The disparity in scores is much greater for *writing* than it is for *math*.

## By Race

### Math

```{r fig.show = "hold", warning= FALSE, message=FALSE, echo=TRUE}
gplot(stest, math, `race/ethnicity`) + geom_density()

gplot(stest, math, `race/ethnicity`) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                              axis.ticks.y = element_blank())
```

An Analysis of Variance would provide more insight into the possible group differences.

```{r message=FALSE, warning=FALSE}
summary(aov(math~`race/ethnicity`, stest))
```

#### Sorted by Food

```{r out.width= "120%"}
stest %>%
  ggplot(aes(y = math, color = `race/ethnicity`)) + geom_boxplot() + 
  theme_bw() + theme(axis.text.x = element_blank(),
                     axis.ticks.x = element_blank()) + 
  facet_wrap(~ lunch)
```

#### Sorted by Test Prep

```{r}
stest %>%
  ggplot(aes(y = math, color = `race/ethnicity`)) + geom_boxplot() + 
  theme_bw() + theme(axis.text.x = element_blank(),
                     axis.ticks.x = element_blank()) + 
  facet_wrap(~ `test prep`)
```


### Writing

```{r fig.show = "hold", message=FALSE, warning=FALSE}
gplot(stest, writing, `race/ethnicity`) + geom_density()

gplot(stest, writing, `race/ethnicity`) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                              axis.ticks.y = element_blank())
```

The difference is less clear with respect to *writing*. Running an ANOVA test here may provide more answers.

```{r}
summary(aov(writing~`race/ethnicity`, stest))
```

## By Parent's Education

### Math

```{r fig.show = "hold", message=FALSE, warning=FALSE}
gplot(stest, math, `parent ed`) + geom_density()

gplot(stest, math, `parent ed`) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                              axis.ticks.y = element_blank())
```


### Writing

```{r fig.show = "hold", message=FALSE, warning=FALSE}
gplot(stest, writing, `race/ethnicity`) + geom_density()

gplot(stest, writing, `race/ethnicity`) + geom_boxplot() + theme(axis.text.y = element_blank(),
                                                              axis.ticks.y = element_blank())
```

# Plotly

```{r}
p <- stest %>%
  ggplot(aes(math, color = `test prep`)) + geom_density() +
  theme_bw() + facet_wrap(~ `race/ethnicity`)
ggplotly(p)
```

# Scrap

### Test prep by Parent education


```{r}
counted <- stest %>%
  group_by(`parent ed`) %>%
  count(`test prep`) %>%
  mutate(n = n/sum(n)) %>%
  arrange(desc(n))
ggplot(counted, aes(n, fct_inorder(`parent ed`), fill = `test prep`)) +
  theme_bw() + geom_col() + ylab("Percent")
```

```{r}
stest %>%
  group_by(`parent ed`) %>%
  count(lunch) %>%
  mutate(n = n/sum(n)) %>%
  ggplot(aes(n, `parent ed`, fill = lunch)) + geom_col() + theme_bw() +
  ylab("Percent")
```

```{r}
stest %>%
  filter(`race/ethnicity` == "group A") %>%
  ggplot(aes(math, color = `test prep`)) + geom_density() +
  theme_bw()

```



